---
title: "DataVisualization"
output: html_document
---

```{r}
library(MASS)
help(Cars93)

data("Cars93", package = "MASS")
#This dataset is about 93 cars on Sale in the USA in 1993. Through making histogram, I would like to see the distribution of the cars' minimum price, maximum price, weight, length, and the price depedning on three types of DriveTrain.

ggplot(Cars93) + geom_histogram(aes(x = Min.Price)) + xlab("Minimum Price")
```

```{r}
ggplot(Cars93) + geom_histogram(aes(x = Max.Price)) + xlab("Maximum Price")
```

```{r}
#Price variables represents average of Min.Price and Max.Price.
ggplot(Cars93) + geom_freqpoly(aes(x = Price))
```

```{r}
ggplot(Cars93) + geom_histogram(aes(x = Length))
```
```{r}
ggplot(Cars93) + geom_histogram(aes(x = Price)) + facet_wrap( ~ DriveTrain)

?facet_wrap
```

# This histogram shows there are a lot more front wheel drive cars in the dataset. Most of the front wheel drive is cheap around $20,000, and has right-skewed distribution, where the mean is bigger than the median. Four wheel drive cars has more symmetrical distribution, and tend to be more cheap. Rear wheel drive cars has a wide range of price, with a few outliers up to $60,000. 

```{r}
#Creating a scatterplot between price and other variables to see the relationship. Variables are including Price, Length, Type of car, Passenger capacity.
library(ggplot2)
Cars93$Type <- as.factor(Cars93$Type)

ggplot(Cars93) + geom_point(aes(x = Length, y = Price, col = Passengers, shape = Type)) + geom_smooth(aes(x = Length, y = Price), colour = "red", se = FALSE)

```

#Cars with longer length tends to have higher price, though not showing a perfect linear relationship. Among the types, small cars tend to have low price. Vans have high passenger capacity but are not really expensive. Midsize car's price has a wide range of price with outliers. Passengers capacity seems to account for relatively little of the variation in price.


```{r}
#Accessing Racial disparities in police interactions with the public, based on Connecticut data. https://openpolicing.stanford.edu/data/

state_patrol <- readRDS(dir(pattern = "rds$"))
library(dplyr)

filter(state_patrol, !is.na(subject_race), !is.na(subject_sex)) %>% group_by(subject_race, subject_sex) %>% summarize(search = mean(search_conducted, na.rm = TRUE), arrest = mean(arrest_made, na.rm = TRUE), either = mean (search_conducted | arrest_made, na.rm = TRUE), both = mean(search_conducted & arrest_made, na.rm = TRUE), n = n())

```

#Except "other", males are more likely to be subjected to be searched by police. Regardless of subject_sex, blacks and hispanics are more likely to be arrested by police, compared to the other races. 

```{r}
#Making boxplots and interpretation
#Data 1) Chicken Weights by Feed Type

#A data frame with 71 observations on the following 2 variables.
#An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens.

data("chickwts")
ggplot(chickwts) + geom_boxplot(aes(x = feed, y = weight))

```

#- Sunflower has the smallest interquartile range, the difference between 75th and 25th percentiles. 

#- The effect the types of feed on chicken weight can be ranked from the highest to the lowest as follows; casein, sunflower, meatmeal, soybean, linseed and horsebean. 

#- Chicken fed by sunflower and horsebean had a smaller range than others, and chicken fed by casein had the largest range. 

#- Looking at medians, the chicken weights with the feed of casein, sunflower, soybean were skewed to right. This indicates that the mean is bigger than median and there are more data that is larger than median. Weight of chicken fed by meatmeal and horsebean is left-skewed, which means that the mean is smaller than the median. 

#- Weight of chicken fed by sunflower had three outliers. Two of them are larger than mean and one of them is smaller than mean. Weight of chicken fed by other types of feed did not have outliers. 

```{r}
data("iris")
#This data set has the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica. I would like to investigate the relationship between Petal Length and the three types of species.

ggplot(iris) + geom_boxplot(aes(x = Species, y = Petal.Length)) + ylab("Petal Length")

```

#Petal lengths of setosa species are the shortest. The highest medians petal length are observed in the case of virginica. Setosas and versicolor shows outliers that is smaller or bigger than the most cases included in each dataset. 

```{r}
data("faithful")
# This dataset measures waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in Yellowstone National Park, Wyoming, USA. 
?faithful
ggplot(faithful) + geom_histogram(aes(x = eruptions)) + xlab("Eruption time in minutes")

```

#The histogram presents a bimodal distribution, with the first peak around 2 minutes and the other around 4.5 minutes. Most of the cases last between four and five minutes, and there are very few explosions that last three minutes and more than five minutes. 



Linear Algebra

< Prove Orthogonality >
## To prove that z is orthogonal to both x and y, both [x]' * z and [y]' * z should be 0.  
## z = Xy = [-x3 * y2 + x2 * y3, x3 * y1 - x1 * y3, -x2 * y1 + x1 * y]'
## [x1, x2, x3] * [-x3 * y2 + x2 * y3, x3 * y1 - x1 * y3, -x2 * y1 + x1 * y2]' = -x1 * x3 * y2 + x1 * x2 * y3 + x2 * x3 * y1 - x1 * x2 * y3 -x2 * x3 * y1 + x1 * x3 * y2 = 0 
## [y1, y2, y3] * [-x3 * y2 + x2 * y3, x3 * y1 - x1 * y3, -x2 * y1 + x1 * y2]' = -x3 * y1 * y2 + x2 * y1 * y3 + x3 * y1 * y2 - x1 * y2 * y3 - x2 * y1 * y3 + x1 * y2 * y3 = 0

```{r}
#< Matrix Calculations >
A <- rbind(c(5, 1, 2), c(6, 2, 3))
B <- matrix(c(3:5, -2, -3, 6), nrow = 2, ncol = 3, byrow = TRUE)
C <- cbind(c(1, -5, -3), c(2, 3, 1))
D <- matrix(c(2, 4, 1, 3), nrow = 2, ncol =2)
```

```{r}
A - B
```

```{r}
A + (5 * B)
```

```{r}
t(B) - C
```

```{r}
D %*% A
```

```{r}
C %*% B
```

